---
title: "Homework6"
output: html_document
date: '2025-02-19'
---

Question 1-3 

We are using a data set collected from an annual survey provided to the Vermont Beekeepers Association to better understand the social demographic of the organization and member satisfaction with aspects of the organization. We aim to better support minority demographics by understanding the satisfaction rating among these groups.

```{r}
ds <- read.csv("Annual Survey Results - Final.csv")

head(ds)


```


Question 4 & 5: Analyzing the data multiple times/ Graphing the data

Instead of changing parameters, we analyzed several questions using different aspects of the data set.

```{r}

library(ggplot2)


ggplot(data = ds, aes(x = factor(Mentorship_program), fill = Gender)) +
  geom_bar(position = "dodge")



```

We used chi-squared tests to calculate the significance of our categorical data.

1) Are the proportion of women that are unsatisfied with the mentorship program significantly different from men?

```{r}

# Create a new column that indicates whether a person is unsatisfied or not
ds$Mentorship_programClean <- ifelse(ds$Mentorship_program %in% c("Unsatisfied", "Very Unsatisfied", "Neutral"),  "Unsatisfied", "Satisfied")

# Create a contingency table for gender vs unsatisfaction status
contingency_table <- table(ds$Gender, ds$Mentorship_programClean)

# Perform the Chi-squared test
chi_squared_test <- chisq.test(contingency_table)

# Print the results of the test
print(chi_squared_test)

# Get the standardized residuals
standardized_residuals <- chi_squared_test$stdres

#Priint residuals
print(standardized_residuals)


print(contingency_table)

```

2) Are beekeepers with more years of beekeeping experience more likely to be satisfied with networking activities?

```{r}

# Create a new column that indicates whether a person is unsatisfied or not
ds$Networking_opportunitiesClean <- ifelse(ds$Networking_opportunities %in% c("Unsatisfied", "Very Unsatisfied", "Neutral"),  "Unsatisfied", "Satisfied")

# Create a contingency table for gender vs unsatisfaction status
contingency_table2 <- table(ds$Beekeeping_experience, ds$Networking_opportunitiesClean)

print(contingency_table2)

# Perform the Chi-squared test
chi_squared_test2 <- chisq.test(contingency_table2)


# Print the results of the test
print(chi_squared_test2)

# Get the standardized residuals
standardized_residuals2 <- chi_squared_test2$stdres

#Priint residuals
print(standardized_residuals2)



```


Question 6-9: For loops

```{r}
library(dplyr)

# list to store results
proportion_results <- list()

# Loop through each gender
for (gender in c("Male", "Female")) {
  # Subset the data for this gender group
  gender_data <- ds[ds$Gender == gender, ]
  
  # list to store proportions for this gender
  gender_proportions <- list()
  
  # Loop through each column
  for (column in c("Annual_meetings", "Mentorship_program", "Access_resources", "Educational_workshops", "Industry_policy_insights", "Networking_opportunities", "News_updates", "Marketing_social")) {
    # Calculate the proportion of each category for this column
    prop_very_satisfied <- mean(gender_data[[column]] == "Very satisfied") * 100
    prop_satisfied <- mean(gender_data[[column]] == "Satisfied") * 100
    prop_unsatisfied <- mean(gender_data[[column]] == "Unsatisfied") * 100
    prop_very_unsatisfied <- mean(gender_data[[column]] == "Very unsatisfied") * 100
    prop_neutral <- mean(gender_data[[column]] == "Neutral") * 100
    prop_noexperience <- mean(gender_data[[column]] == "No prior experience") * 100
    
    # Store the results in the list
    gender_proportions[[column]] <- c("Very satisfied" = prop_very_satisfied,
                                      "Satisfied" = prop_satisfied,
                                      "Unsatisfied" = prop_unsatisfied,
                                      "Very unsatisfied" = prop_very_unsatisfied,
                                      "Neutral" = prop_neutral,
                                      "No prior expreience" = prop_noexperience)
  }
  
  # Store the results for this gender
  proportion_results[[gender]] <- gender_proportions
}

# Print the results
print(proportion_results)

  
```








